### 2017-10-14 writing/paper/sections/coincidences.tex
Final edits from Murray
Fix minor speling mistakes "patter"->"pattern"
"leftmostpoint"->"leftmost point"

Reflow text to line length 80

In order to describe the rules it is useful to have a notion for inserting points, ascents, and
descents into a mesh pattern.
\begin{definition}
\label{def:ap}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\). We define
a mesh pattern \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a
point is \emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new underlying
classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+1 & \text{if } k = i+1\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+1 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-1) & \text{if } \tau(k)\le j \text{ and }k> i+1\\
    \tau(k-1)+1 & \text{if } \tau(k)> j \text{ and }k> i+1\\
\end{cases}
\end{equation*}
While the mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-1}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-1,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-1,\ell-1}\in R\}\\
\end{aligned}
\end{equation*}
\end{definition}
In addition, we give the following definitions:

\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that \(\boks{i,j}\notin R\) and
\(p^\boks{i,j} = (\tau^\prime,R^\prime)\) is as defined in \DefinitionRef{def:ap}. We define
the following four modifications of \(p^\boks{i,j}\).
\begin{align*}
p^\boksup{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j+1},\boks{i+1,j+1}\})\\
p^\boksright{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i+1,j},\boks{i+1,j+1}\})\\
p^\boksdown{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i+1,j}\})\\
p^\boksleft{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i,j+1}\})\\
\end{align*}
\end{definition}
Informally, these are considering the topmost, rightmost, leftmost, or bottommost point in
\(\boks{i,j}\). We collect the resulting mesh patterns in a set
\begin{equation*}
p^\boksall{i,j}=\left\{
p^\boks{i,j},
p^\boksright{i,j},
p^\boksleft{i,j},
p^\boksup{i,j},
p^\boksdown{i,j}\right\}
\end{equation*}
See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.

@@

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \(\boks{1,0}\) is empty then we have an occurrence of \(m_2\). Now look at
the case when this region is not empty, and consider choosing the rightmost
point in the region.
This gives us an occurrence of the following mesh pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of First Dominating Patter Rule we obtain the following
mesh pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
The highlighted points are an occurrence
of the mesh pattern \(m_2\). This gives rise to the following
rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), then
    \(m_1 \coinc{\pi} m_2\) if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \(m_1^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(m_2\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\) as before.
    Now consider a permutation \(\omega \in \av{\pi}\) that contains an
    occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\)
    is empty then we have an occurrence of \(m_2\).
    By condition~\ref{prop:dom3:condocc} of the proposition there exists
    a pattern \(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some
    pattern \(r'\) coincident with \(r\). If the region is non-empty then there exists
    an occurrence of \(r\) in \(\omega\). Therefore there is an occurrence of \(r'\)
    in \(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
    Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the
    two patterns are coincident.
\end{proof}

@@

\subsection{Coincidence classes of Av(\{321, (12, \textit{R})\}).}
When considering coincidences of mesh patterns with underlying classical
pattern \(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
established rules gives no coincidences. Through experimentation we discover
that there are \(7\) \(\coinc[comp]{321}\)-equivalence classes are unexplained.
Since the number of coincidences is
so small we will reason for these coincidences without attempting to generalise
into concrete rules.

@@

The patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \text { and } m_2 \pattern{}{1,2}{0/0, 0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are \(\coinc[comp]{321}\)-equivalent. (There are 3 symmetries of these
patterns that are also equivalent to each other by the same reasoning.)

@@

In order to prove this coincidence we will proceed by mathematical
induction on the number of points in the region corresponding to the middle box.
We call this number \(n\).
\begin{description}
    \item [Base Case \((n=0)\)] The base case holds since we can freely shade the
        box if it contains no points.
    \item [Inductive Hypothesis \((n=k)\)] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.
    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, giving the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty then
        we have an occurrence of \(m_2\) and are done. If this box contains
        any points then we gain some extra shading on the mesh pattern
        as any points in the boxes \(\boks{1,2}\) and \(\boks{1,3}\) would create an occurrence of the dominating pattern \(\perm{3,2,1}\)
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.
\end{description}
By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
\(m_1\) so the two patterns are coincident.
This argument applies to another two pairs of classes. Therefore in
total in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\) there are \(\coinc{321}\)-equivalence
classes.

>>>>

In order to describe the rules it is useful to have a notion for inserting
points, ascents, and descents into a mesh pattern.
\begin{definition}
\label{def:ap}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that
\(\boks{i,j}\notin R\). We define a mesh pattern \(p^\boks{i,j} =
(\tau^\prime,R^\prime)\) of length \(n+1\) as the pattern where a point is
\emph{inserted} into the box \(\boks{i,j}\) in \(G(p)\). Formally the new
underlying classical pattern is defined by
\begin{equation*}
\tau^\prime(k) = \begin{cases}
    j+1 & \text{if } k = i+1\\
    \tau(k) & \text{if } \tau(k)\le j \text{ and }k\le i\\
    \tau(k)+1 & \text{if } \tau(k)> j \text{ and }k\le i\\
    \tau(k-1) & \text{if } \tau(k)\le j \text{ and }k> i+1\\
    \tau(k-1)+1 & \text{if } \tau(k)> j \text{ and }k> i+1\\
\end{cases}
\end{equation*}
While the mesh becomes
\begin{equation*}
\begin{aligned}
R^\prime=&\{\boks{k,\ell}\mid k\le i, \ell\le j, \boks{k,\ell}\in R\} \cup\\
&\{\boks{k,\ell}\mid k\le i, \ell > j, \boks{k,\ell-1}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell \le j, \boks{k-1,\ell}\in R\}\cup\\
&\{\boks{k,\ell}\mid k > i, \ell > j, \boks{k-1,\ell-1}\in R\}\\
\end{aligned}
\end{equation*}
\end{definition}
In addition, we give the following definitions:
\begin{definition}
Let \(p=(\tau,R)\) be a mesh pattern of length \(n\) such that
\(\boks{i,j}\notin R\) and \(p^\boks{i,j} = (\tau^\prime,R^\prime)\) is as
defined in \DefinitionRef{def:ap}. We define the following four modifications of
\(p^\boks{i,j}\).
\begin{align*}
p^\boksup{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j+1},\boks{i+1,j+1}\})\\
p^\boksright{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i+1,j},\boks{i+1,j+1}\})\\
p^\boksdown{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i+1,j}\})\\
p^\boksleft{i,j} &= (\tau^\prime, R^\prime \cup \{\boks{i,j},\boks{i,j+1}\})\\
\end{align*}
\end{definition}
Informally, these are considering the topmost, rightmost, leftmost, or
bottommost point in \(\boks{i,j}\). We collect the resulting mesh patterns in a
set
\begin{equation*}
p^\boksall{i,j}=\left\{
p^\boks{i,j},
p^\boksright{i,j},
p^\boksleft{i,j},
p^\boksup{i,j},
p^\boksdown{i,j}\right\}
\end{equation*}
See \FigureRef{fig:addp} for an example of adding a point into a mesh pattern.

@@

Consider an occurrence of \(m_1\) in a permutation, if the region corresponding
to the box \(\boks{1,0}\) is empty then we have an occurrence of \(m_2\). Now
look at the case when this region is not empty, and consider choosing the
rightmost point in the region. This gives us an occurrence of the following mesh
pattern.
\begin{equation*}
    \pattern{}{2,1,3}{0/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
By application of First Dominating Pattern Rule we obtain the following mesh
pattern
\begin{equation*}
    \pattern{2,3}{2,1,3}{0/3,1/3,2/0,2/1,3/0,3/1,3/2}
\end{equation*}
The highlighted points are an occurrence of the mesh pattern \(m_2\). This gives
rise to the following rule:

\begin{proposition}[Third Dominating Pattern Rule]
    \label{prop:dom3}
    Given two mesh patterns \(m_1 =(\sigma, R_1)\) and \(m_2 = (\sigma, R_2)\),
    and a dominating classical pattern \(\pi = (\pi,\emptyset)\), then
    \(m_1 \coinc{\pi} m_2\) if
    \begin{enumerate}
        \item The mesh set \(R_2 = R_1 \cup \{\boks{a,b}\}\)
        \item\label{prop:dom3:condocc} One of the patterns in \(m_1^\boksall{a,b}\)
            is coincident with a mesh pattern containing an occurrence of
            \(m_2\) as a subpattern.
    \end{enumerate}
\end{proposition}
\begin{proof}
    \NoteRef{not:downcmesh} implies \(\av{\{\pi,m_1\}} \subseteq \av{\{\pi,m_2\}}\)
as before. Now consider a permutation \(\omega \in \av{\pi}\) that contains an
occurrence of \(m_1\). If the region corresponding to the box \(\boks{a,b}\) is
empty then we have an occurrence of \(m_2\). By
condition~\ref{prop:dom3:condocc} of the proposition there exists a pattern
\(r\) in \(m_1^\boksall{a,b}\) such that \(m_2\) is a subpattern in some pattern
\(r'\) coincident with \(r\). If the region is non-empty then there exists an
occurrence of \(r\) in \(\omega\). Therefore, there is an occurrence of \(r'\) in
\(\omega\), which implies that there is an occurrence of \(m_2\) in \(\omega\).
Thus \(\av{\{\pi,m_2\}} \subseteq \av{\{\pi,m_1\}}\) and the two patterns are
coincident.
\end{proof}

@@

\subsection{Coincidence classes of Av(\{321, (12, \textit{R})\}).}
When considering coincidences of mesh patterns with underlying classical pattern
\(\perm{1,2}\) in \(\av{\perm{3,2,1}}\) application of the previously
established rules gives no coincidences. Through experimentation we discover
that there are \(7\) \(\coinc[comp]{321}\)-equivalence classes that are
unexplained. Since the number of coincidences is so small we will reason for
these coincidences without attempting to generalise into concrete rules.

@@

The patterns
\begin{equation*}
    m_1 = \pattern{}{1,2}{0/0, 0/2, 1/1, 1/2, 2/1} \et m_2 \pattern{}{1,2}{0/0,
    0/1, 0/2, 1/1, 1/2, 2/1}
\end{equation*}
are \(\coinc[comp]{321}\)-equivalent. (There are 3 symmetries of these
patterns that are also equivalent to each other by the same reasoning.)

@@

In order to prove this coincidence we will proceed by mathematical
induction on the number of points in the region corresponding to the middle box.
We call this number \(n\).
\begin{description}
    \item [Base Case \((n=0)\)] The base case holds since we can freely shade the
        box if it contains no points.
    \item [Inductive Hypothesis \((n=k)\)] Suppose that we can find an occurrence
        of the second pattern if we have an occurrence of the first with \(k\) points
        in the middle box.
    \item [Inductive Step \((n=k+1)\)] Suppose that we have \((k+1)\) points in
        the middle box. Choose the bottom most point in the middle box, giving the
        mesh pattern
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{}{1,2,3}{0/1,0/2,0/3,1/1,2/1,3/0,3/1,3/2}
            \draw (2.5,0.5) node {\(X\)};
        \end{tikzpicture}
    \end{equation*}
        Now we need to consider the box labelled \(X\). If this box is empty
        then we have an occurrence of \(m_2\) and are done. If this box contains
        any points then we gain some extra shading on the mesh pattern as any
        points in the boxes \(\boks{1,2}\) and \(\boks{1,3}\) would create an
        occurrence of the dominating pattern \(\perm{3,2,1}\)
     \begin{equation*}
        \begin{tikzpicture}[scale=0.5]
            \modpattern[5]{2,3}{1,2,3}{0/1,0/2,0/3,1/1,1/2,1/3,2/1,3/0,3/1,3/2}
            \filldraw (2.5,0.5) circle (4 pt);
        \end{tikzpicture}
    \end{equation*}
        The two highlighted points form an occurrence of \(m_1\) with \(k\)
        points in the middle box, and thus by the Inductive Hypothesis we
        are done.
\end{description}
By induction we have that every occurrence of \(m_1\) leads to an occurrence of \(m_2\)
and by \NoteRef{not:downcmesh} every occurrence of \(m_2\) is an occurrence of
\(m_1\) so the two patterns are coincident.
This argument applies to another two pairs of classes. Therefore we have
explained all 7 of the coincidences in \(\av{\{\perm{3,2,1}, \mperm{1,2}{R}\}}\)
and there are 213 \(\coinc{321}\)-equivalence classes.
